---
title: "Supplemental Materials"
output: 
  bookdown::pdf_document2:
    toc: false
bibliography: bibliography.bib
---

# Methods

Contact call similarity was measured using three different methods:

1) dynamic timewarping on fundamental frequency traces from Luscinia (DTW): see main text

2) cross correlation on fundamental frequency traces from Luscinia (CC): a custom written script (see ANALYSIS/CODE/function/simple.cc.R) was used to slide two traces over each other and measure the pointwise absolute difference. The minimal value (maximal overlap) was standardised by the length of the longest trace and used as similarity measure. 

3) spectrographic cross correlation (SPCC): for each call we applied a high-pass filter using *ffilter* from the package *seewave* [@seewave] with settings: from = 500 Hz. A custom written script was then used to create spectrograms (see ANALYSIS/CODE/function/cutte.spectro.R). We used *specgram* from the package *signal* [@signal] with window = 512 and overlap = 450 to create the basic spectrogram. Only the 1-6 kHz range was included in the final spectrogram. All pixels were standardised. All pixels with a value lower than 1.3 were then set to 0 and pixels with a value higher than 1.8 were set to 1.8. This was done to remove as much noise as possible and reduce the effect of particularly loud sections of the call. As a final stepped each pixel was divided by the summed value of all pixels, making sure that each spectrogram summed to one. This was done to make long and short call comparable. 

Dimensionality of the resulting distance matrices was reduced to two using three methods: 

1) principle coordinate analysis (PCO): see main text 

2) principle component analysis (PCA): using *princomp* from the basic *stats* package in R

3) uniform manifold approximation and projection (UMAP): using *umap* from the package *umap* [@umap] with the settings: input = 'dist', n_neighbors = 10, spread = 1, min_dist = 0.1

For all possible combinations of similarity measures with dimension reduction we ran the same Bayesian multilevel model on the two remaining dimensions (see main text).

# Results

Figure 1-9 show the composit figures for all methods.

```{r cc-pca, fig.cap="Result for CC - PCA. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - cc - pca.pdf')
```

```{r cc-pco, fig.cap="Result for CC - PCO. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - cc - pco.pdf')
```

```{r cc-umap, fig.cap="Result for CC - UMAP. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - cc - umap.pdf')
```

```{r dtw-pca, fig.cap="Result for DTW - PCA. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - dtw - pca.pdf')
```

```{r dtw-pco, fig.cap="Result for DTW - PCO. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - dtw - pco.pdf')
```

```{r dtw-umap, fig.cap="Result for DTW - UMAP. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure luscinia - dtw - umap.pdf')
```

```{r spcc-pca, fig.cap="Result for SPCC - PCA. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure spcc - pca.pdf')
```

```{r spcc-pco, fig.cap="Result for SPCC - PCO. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure spcc - pco.pdf')
```

```{r spcc-umap, fig.cap="Result for SPCC - UMAP. Colour represent city (see legend). Clock-wise: City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 1; Scatter-plot of all calls included in the model; City (thick) and park (thin) averages (dots) and 50, 90 and 95% intervals for DIM 2.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/composite figure spcc - umap.pdf')
```

\newpage

Figure 10-18 show the sigma values for all methods. 

```{r sigmascc-pca, fig.cap="Result for CC - PCA. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - cc - pca.pdf')
```

```{r sigmascc-pco, fig.cap="Result for CC - PCO. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - cc - pco.pdf')
```

```{r sigmascc-umap, fig.cap="Result for CC - UMAP. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - cc - umap.pdf')
```

```{r sigmasdtw-pca, fig.cap="Result for DTW - PCA. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - dtw - pca.pdf')
```

```{r sigmasdtw-pco, fig.cap="Result for DTW - PCO. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - dtw - pco.pdf')
```

```{r sigmasdtw-umap, fig.cap="Result for DTW - UMAP. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas luscinia - dtw - umap.pdf')
```

```{r sigmasspcc-pca, fig.cap="Result for SPCC - PCA. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas spcc - pca.pdf')
```

```{r sigmasspcc-pco, fig.cap="Result for SPCC - PCO. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas spcc - pco.pdf')
```

```{r sigmasspcc-umap, fig.cap="Result for SPCC - UMAP. Sigma values for city and park.", echo=FALSE}
knitr::include_graphics('../../RESULTS/figures/sigmas spcc - umap.pdf')
```

\newpage

Figure 19 shows the trade history for the four countries for which we have recordings. 

```{r trade, fig.cap="Trade in monk parakeets through time recorded by CITES.", echo=FALSE}
knitr::include_graphics('../../RESULTS/CITES trade/trade graphs.pdf')
```

# References